---
title: "Trial 1"
author: "Levia"
output: html_document
date: "2023-04-03"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

This is a trial for R Markdown 
Today is 4 April 2023

Triple ' followed by {r} will start an r block
e.g. Loading library  
all library used in this project
``` {r, warning = FALSE, message = FALSE}
library(actuar)
library(cluster)
library(dplyr)
library(EnvStats)
library(factoextra)
library(fitdistrplus)
library(ggplot2)
library(goftest)
library(MASS)
library(randomForest)
library(readr)
library(readxl)
library(rmarkdown)
library(stringi)
library(stringr)
library(tibble)
library(tidyr)
library(tidyverse)
```

e.g. loading data
``` {r}
load(file = "census.RData")
load(file = "emissions.RData")
load(file = "hazards.RData")
load(file = "inflation.RData")
load(file = "rates.RData")
```

e.g. glimpse
```{r}
glimpse(inflation)
```

### OK LOADING DONE
note to self:  
original data was csv which cant be read somehow here,  
so do read.csv of each files according to each form.  
then do save each files to .RData form. 
then load it to the rmd space.  
[RMarkdown Cheat Sheet](https://www.rstudio.com/wp-content/uploads/2015/02/rmarkdown-cheatsheet.pdf)
  
## NEXT seperate the EDA, Fitting, and Simulation




# 1. Prelim_Clustering ####
Initially we try to cluster the hazards based on it's event and the damages it brought.  
```{r, warning=FALSE, message=FALSE}
# Setting up the environment, load the necessary packages
library(cluster)
library(dplyr)
library(factoextra)
library(ggplot2)
library(readr)
library(stringi)
library(tidyverse)
```

**Load the data**
```{r, eval=TRUE} 
load(file = "hazards.RData")
glimpse(hazards)
```
**Data Cleaning**  

* Change the `Region` column from integer to character  
  
* Filter event that has impact, identified from:  
  
  + Property damage > 0 or  
    
  + Injuries > 0 or  
    
  + Fatalities > 0  
```{r}
hazards$Region = as.character(hazards$Region)
hazards = hazards %>%
  filter(hazards$Property.Damage > 0 | hazards$Injuries > 0 | hazards$Fatalities > 0)
glimpse(hazards)
head(hazards, 5)
```
_Additional Steps_
```{r}
# combine anything with flooding & costal 
# combine storm/hail/wind/lightening as they represent extreme weather conditions cause meteorological hazards.
for (i in 1:nrow(hazards)) {
  if (stri_detect_fixed(hazards[i,"Hazard.Event"], "Hurricane") == TRUE ) {
    hazards[i,"Hazard.Event"] = "Hurricane" 
  } else if ( stri_detect_fixed(hazards[i,"Hazard.Event"], "Coastal") == TRUE ) {
    hazards[i,"Hazard.Event"] = "Coastal" 
  } else if (stri_detect_fixed((hazards[i,"Hazard.Event"]), "Flooding") == TRUE){
    hazards[i,"Hazard.Event"] = "Flooding" 
  } else if (stri_detect_fixed((hazards[i,"Hazard.Event"]), "Storm") == TRUE || 
             stri_detect_fixed((hazards[i,"Hazard.Event"]), "Hail") == TRUE ||
             stri_detect_fixed((hazards[i,"Hazard.Event"]), "Wind") == TRUE ||
             stri_detect_fixed((hazards[i,"Hazard.Event"]), "Lightning") == TRUE  ||
             stri_detect_fixed((hazards[i,"Hazard.Event"]), "Winter Weather") == TRUE){
    hazards[i,"Hazard.Event"] = "Severe Weather" 
  } else if (stri_detect_fixed((hazards[i,"Hazard.Event"]), "Drought") == TRUE || 
             stri_detect_fixed((hazards[i,"Hazard.Event"]), "Heat") == TRUE) {
             hazards[i,"Hazard.Event"] = "Drought/Heat" 
  }
}

# removed year and region columns to evaluate the historical event's clustering regardless of time and place
hazards.agg <- hazards %>%
  dplyr::select(-c(Year, Region))

hazards.agg2 =  hazards.agg %>% 
  group_by(Hazard.Event) %>% 
  summarize(Property.Damage.agg = sum(Property.Damage), 
         Injuries.agg = sum(Injuries),Fatalities.agg = sum(Fatalities),
         Freq = n()) %>%
  mutate(Property.Damage.avg = Property.Damage.agg/Freq , 
         Injuries.avg = Injuries.agg/Freq,Fatalities.avg =Fatalities.agg/Freq) %>%
  remove_rownames %>% column_to_rownames(var="Hazard.Event")

# drop fog/landslide due to its scarcity 
hazards.agg2 = hazards.agg2[!(row.names(hazards.agg2) %in% c("Fog","Landslide")),]
```
**Event's clustering via kmeans**
```{r}
df = scale(hazards.agg2)
k4 <- kmeans(df[,c("Property.Damage.avg", "Fatalities.avg","Injuries.avg","Freq")], centers = 3)

distance <- get_dist(df)
```
```{r, echo=FALSE}
fviz_cluster(k4, df, geom = "text")
```





# 2. Goodness_of_Fit_Tests ####
```{r, warning=FALSE, message=FALSE}
# Setting up the environment, load the necessary packages
library(actuar)
library(dplyr)
library(EnvStats)
library(fitdistrplus)
library(kableExtra)
library(knitr)
library(ggplot2)
library(goftest)
library(MASS)
library(tidyr)
```
**Setting up and data preparation**
```{r, warning=FALSE}
options(scipen=999)
set.seed(123)

# load data
load(file = "census.RData")
load(file = "hazards.RData")

# filtering non-impacting hazards
hazards <- hazards %>% arrange(Year) %>% filter(Property.Damage != 0)
# convert census type such that it could be read
census[,] <- sapply(census[, ], as.numeric) 
str(census)

# create clustering
clusters <- hazards %>%
  mutate(Hazard.Event = 
           recode(Hazard.Event, 
                  "Hail/ Wind" = "Severe Weather",
                  "Winter Weather"= "Severe Weather",
                  "Wind" = "Severe Weather", 
                  "Hail/ Severe Storm/Thunder Storm" = "Severe Weather",
                  "Severe Storm/Thunder Storm" = "Severe Weather", 
                  "Hail" = "Severe Weather", 
                  "Severe Storm/Thunder Storm/ Wind" = "Severe Weather",
                  "Lightning" = "Severe Weather",
                  "Lightning/ Wind" = "Severe Weather",
                  "Tornado"  = "Tornado",
                  "Coastal/ Wind" = "Severe Weather",
                  "Hurricane/Tropical Storm" = "Hurricane",
                  "Flooding" = "Flooding",
                  "Hail/ Lightning/ Wind" = "Severe Weather",
                  "Hail/ Severe Storm/Thunder Storm/ Wind" = "Severe Weather",
                  "Lightning/ Severe Storm/Thunder Storm" = "Severe Weather",
                  "Wildfire" = "Wildfire",
                  "Hail/ Lightning/ Severe Storm/Thunder Storm" = "Severe Weather",
                  "Hail/ Lightning/ Severe Storm/Thunder Storm/ Wind" = "Severe Weather",
                  "Tornado/ Wind" = "Tornado",
                  "Hail/ Tornado/ Wind" = "Tornado",
                  "Lightning/ Severe Storm/Thunder Storm/ Wind" = "Severe Weather",
                  "Severe Storm/Thunder Storm/ Winter Weather" = "Severe Weather",
                  "Wind/ Winter Weather" = "Severe Weather",
                  "Coastal/ Hurricane/Tropical Storm/ Wind" = "Hurricane",
                  "Severe Storm/Thunder Storm/ Wind/ Winter Weather" = "Severe Weather",
                  "Flooding/ Severe Storm/Thunder Storm" = "Flooding",
                  "Coastal" = "Coastal",
                  "Hail/ Lightning" = "Severe Weather",
                  "Flooding/ Lightning/ Severe Storm/Thunder Storm" = "Flooding",
                  "Heat" = "Drought/ Heat",
                  "Drought/ Heat" = "Drought/ Heat",
                  "Flooding/ Severe Storm/Thunder Storm/ Wind" = "Flooding",
                  "Flooding/ Wind" = "Flooding",
                  "Drought" = "Drought/ Heat",
                  "Hurricane/Tropical Storm/ Severe Storm/Thunder Storm" = "Hurricane",
                  "Coastal/ Flooding" = "Coastal",
                  "Flooding/ Lightning" = "Flooding",
                  "Coastal/ Severe Storm/Thunder Storm/ Wind" = "Coastal",
                  "Coastal/ Severe Storm/Thunder Storm" = "Coastal",
                  "Coastal/ Flooding/ Severe Storm/Thunder Storm/ Wind" = "Coastal",
                  "Lightning/ Tornado/ Wind" = "Tornado",
                  "Flooding/ Lightning/ Wind" = "Flooding",
                  "Flooding/ Hail"= "Flooding",
                  "Hail/ Tornado" = "Tornado",
                  "Hail/ Severe Storm/Thunder Storm/ Wind/ Winter Weather" = "Severe Weather",
                  "Coastal/ Hurricane/Tropical Storm/ Severe Storm/Thunder Storm/ Wind" = "Hurricane",
                  "Flooding/ Hail/ Wind"= "Flooding",
                  "Fog" = "Fog",
                  "Severe Storm/Thunder Storm - Wind" = "Severe Weather")) %>%
  filter(Hazard.Event != "Fog")

summary(clusters)
```
  
_Identified Hazard Event based on the clustering_  
```{r, echo=FALSE}
(kable(unique(clusters$Hazard.Event), "pipe", col.names = "Hazard Events", align = "c"))
```
### Hazard Analysis
##### Frequency by hazard to determine the level of hazard
```{r}
frequency_hazard <- clusters %>% 
  group_by(clusters$Hazard.Event) %>% 
  summarise(events.pa = n(), prop_damage = sum(Property.Damage),
            damage_per_event = sum(Property.Damage)/n(), risk_level =
              ifelse(damage_per_event <= 1000000,"minor",
                     ifelse(damage_per_event<= 10000000,"medium","major")))

## find the mean of losses for each risk level
minor_mean <- frequency_hazard %>% 
    group_by(risk_level) %>% 
    filter(risk_level == "minor") %>% 
    summarise(dam = mean(damage_per_event)) #517881

medium_mean <- frequency_hazard %>% 
  group_by(risk_level) %>% 
  filter(risk_level == "medium") %>% 
  summarise(dam = mean(damage_per_event)) #4345546

major_mean <- frequency_hazard %>% 
  group_by(risk_level) %>% 
  filter(risk_level == "major") %>% 
  summarise(dam = mean(damage_per_event)) #134796018

## assign weighting for each risk
damage_table <- rbind(minor_mean, medium_mean, major_mean)
colnames(damage_table) <- c("Risk Level", "Average Damage")
kable(damage_table, align = "cr") %>%
  kable_styling(full_width = FALSE)

weighting_minor <- as.numeric(damage_table[1,2]/sum(damage_table[,2]))
weighting_medium <- as.numeric(damage_table[2,2]/sum(damage_table[,2]))
weighting_major <- as.numeric(damage_table[3,2]/sum(damage_table[,2]))
## check
sum(weighting_minor+weighting_medium+weighting_major)
```
##### Frequency by Region
```{r}
frequency_region <- clusters %>% 
  group_by(Region,Year) %>% 
  dplyr::summarise(tot_minor = sum(Hazard.Event == "Severe Weather"),
                   tot_medium = (sum(Hazard.Event == "Coastal")+
                                   sum( Hazard.Event == "Drought/ Heat")+
                                   sum(Hazard.Event == "Flooding")+
                                   sum(Hazard.Event == "Tornado")),
                   tot_major = (sum(Hazard.Event == "Wildfire")+
                                  sum( Hazard.Event == "Hurricane")),
                   tot_event = tot_minor + tot_medium + tot_major,
                   prop_damage = sum(Property.Damage), .groups='drop')

## quantify the region to low risk, med risk and high risk
# checking per region the no of minor, med and major risk and the prop damage
analyse_reg1 <- frequency_region %>% 
  filter(Region==1) %>% 
  summarise(Region = "Region 1", 
            minor = sum(tot_minor), medium = sum(tot_medium),
            major = sum(tot_major), sum_damage = sum(prop_damage))

analyse_reg2 <- frequency_region %>% 
  filter(Region==2) %>% 
  summarise(Region ="Region 2",
            minor = sum(tot_minor), medium = sum(tot_medium),
            major = sum(tot_major), sum_damage = sum(prop_damage))

analyse_reg3 <- frequency_region %>% 
  filter(Region==3) %>% 
  summarise(Region ="Region 3",
            minor = sum(tot_minor), medium = sum(tot_medium),
            major = sum(tot_major), sum_damage = sum(prop_damage))

analyse_reg4 <- frequency_region %>% 
  filter(Region==4) %>% 
  summarise(Region ="Region 4",
            minor = sum(tot_minor), medium = sum(tot_medium),
            major = sum(tot_major), sum_damage = sum(prop_damage))

analyse_reg5 <- frequency_region %>% 
  filter(Region==5) %>% 
  summarise(Region ="Region 5",
            minor = sum(tot_minor), medium = sum(tot_medium),
            major = sum(tot_major), sum_damage = sum(prop_damage))

analyse_reg6 <- frequency_region %>% 
  filter(Region==6) %>% 
  summarise(Region ="Region 6",
            minor = sum(tot_minor), medium = sum(tot_medium),
            major = sum(tot_major), sum_damage = sum(prop_damage))

region_analysis <- rbind(analyse_reg1,analyse_reg2,analyse_reg3,analyse_reg4,analyse_reg5,analyse_reg6)
colnames(region_analysis)[5] <- "Total Damage"
kable(region_analysis) %>% 
  kable_styling(full_width = F)

# find the mean and median for each classes of event per region
minor_med <- median(region_analysis$minor) #379
medium_med <- median(region_analysis$medium) #117.5
major_med <- median(region_analysis$major) #11

minor_mean <- mean(region_analysis$minor) #369.833
medium_mean <- mean(region_analysis$medium) #121
major_mean <- mean(region_analysis$major) #14.333

damage_r1 <- region_analysis %>% filter(Region == "Region 1") %>%
  summarise(damage = minor*weighting_minor + medium*weighting_medium +
              major*weighting_major)
damage_r2 <- region_analysis %>% filter(Region == "Region 2") %>%
  summarise(damage = minor*weighting_minor + medium*weighting_medium +
              major*weighting_major)
damage_r3 <- region_analysis %>% filter(Region == "Region 3") %>%
  summarise(damage = minor*weighting_minor + medium*weighting_medium +
              major*weighting_major)
damage_r4 <- region_analysis %>% filter(Region == "Region 4") %>%
  summarise(damage = minor*weighting_minor + medium*weighting_medium +
              major*weighting_major)
damage_r5 <- region_analysis %>% filter(Region == "Region 5") %>%
  summarise(damage = minor*weighting_minor + medium*weighting_medium +
              major*weighting_major)
damage_r6 <- region_analysis %>% filter(Region == "Region 6") %>%
  summarise(damage = minor*weighting_minor + medium*weighting_medium +
              major*weighting_major)

reg_damage_analysis <- rbind(damage_r1, damage_r2, damage_r3, damage_r4, damage_r5, damage_r6)

reg_damage_analysis <- tibble::rownames_to_column(reg_damage_analysis, "Region")
colnames(reg_damage_analysis)[2] <- "Adj. Freq"

kable(reg_damage_analysis, align = "cr", caption = "Weight Adjusted Hazard's Frequency") %>% 
  kable_styling(full_width = F)

quantile(reg_damage_analysis$`Adj. Freq`,c(.33, .67))
#     33%      67% 
# 14.74977 16.65021 

region_risk <- ifelse(reg_damage_analysis$`Adj. Freq` <=
                        quantile(reg_damage_analysis$`Adj. Freq`,c(.33)), "Low Risk",
                ifelse(reg_damage_analysis$`Adj. Freq` <=
                        quantile(reg_damage_analysis$`Adj. Freq`,c(.67)), "Medium Risk","High Risk"))

region_risk <- data.frame(region_risk)

region_risk <- tibble::rownames_to_column(region_risk, "Region")
colnames(region_risk)[2] <- "Risk Level"

kable(region_risk, align = "cl", caption = "Region Risk Level") %>%
  kable_styling(full_width = F)

```                     
  
### Census Data Analysis
go to [this link](/Trial-2)
